### 2023-10-25 src/scripts/distance/9-distance-posthoc-analysis.Rmd
added posthoc analysis of distance modelling stuff for paper

---
title: "9-distance-posthoc-analysis"
author: "Brandon Edwards"
date: "2023-10-24"
output: html_document
---

```{r setup, include=FALSE}
library(magrittr)
knitr::opts_chunk$set(echo = TRUE)
```
```{r read_data, echo = FALSE}
####### Read Data #################################

tau_df <- read.csv("../../../data/generated/tau.csv")
tau_diff <- read.csv("../../../data/generated/tau_differences.csv")
lppd_pointwise <- read.csv("../../../data/generated/lppd_distance.csv")
lppd_species <- read.csv("../../../data/generated/lppd_summary_distance.csv")
params <- read.csv("../../../data/generated/distance_params.csv")
traits <- read.csv("../../../data/raw/traits.csv")

ibp_codes <- read.csv("../../../data/raw/IBP-Alpha-Codes20.csv")
families <- read.csv("../../../data/raw/NACC_list_species.csv")

####### Set Constants #############################

sp_pred <- c("LCTH", "LEPC", "HASP", "SPOW", "KIWA", "BITH")
```

# Cross Validation Statistics

Table of Model Preference from Cross-validation:
```{r mod_pref, echo = FALSE}
print(data.frame(table(lppd_species$Model_Preference)))
```

Mean LPPD Difference overall: `r mean(lppd_pointwise$Difference)`

"Sample Size/Model Selection of Species with >20% difference in EDR:
```{r n_mod_selection, echo = FALSE}
tau_diff_pred <- merge(tau_diff, lppd_species[, c("Species", "Mean_Difference", "Model_Preference")],
                       by = "Species")
print(tau_diff_pred[which(tau_diff_pred$Label != ""), ])
```

```{r, echo = FALSE}
l25 <- quantile(tau_diff$Single)[2]
```
25th quantile of EDR in single-species model: `r l25`
```{r, echo = FALSE}
sp_l25 <- tau_diff_pred[which(tau_diff_pred$Single <= l25),]
```

Number of species with single-species EDR at/below 25th quantile: `r nrow(sp_l25)`

Number of species at/below 25th quantile that selected multi-species: `r length(which(sp_l25$Model_Preference == "MS"))` (`r length(which(sp_l25$Model_Preference == "MS")) / nrow(sp_l25)`%)

# Parameter Statistics
```{r, echo = FALSE}
print(params[, c("variable", "mean", "median", "q5", "q95")])
```

```{r, echo = FALSE}
traits_red <- traits[which(traits$Code %in% tau_diff$Species), ]
```

Number of resident species: `r nrow(traits_red) - sum(traits_red$Migrant)`

Number of migrant species: `r sum(traits_red$Migrant)`

# Predictions
```{r, echo = FALSE}
fam_df <- data.frame(Species = tau_df$Code, Tau = exp(tau_df$mean) * 100) %>%
  merge(ibp_codes[, c("SPEC", "COMMONNAME")],
        by.x = "Species", by.y = "SPEC") %>%
  merge(families[, c("common_name", "family")],
        by.x = "COMMONNAME", by.y = "common_name") %>%
  merge(traits[, c("Code", "Mass", "Pitch")],
        by.x = "Species", by.y = "Code")

fam_df_sparrow <- fam_df[which(fam_df$family == "Passerellidae"), ]
fam_df_sparrow <- fam_df_sparrow[-which(fam_df_sparrow$Species == "HASP"), ]

fam_df_thrasher <- fam_df[which(fam_df$family == "Mimidae"), ]
fam_df_thrasher <- fam_df_thrasher[-which(fam_df_thrasher$Species == "LCTH"), ]

fam_df_warbler <- fam_df[which(fam_df$family == "Parulidae"), ]
fam_df_warbler <- fam_df_warbler[-which(fam_df_warbler$Species == "KIWA"), ]

fam_df_thrush <- fam_df[which(fam_df$family == "Turdidae"), ]
fam_df_thrush <- fam_df_thrush[-which(fam_df_thrush$Species == "BITH"), ]
```

Estimates of EDR for predicted species:
```{r cr_pred, echo = FALSE}
tau_pred <- tau_df[which(tau_df$Code %in% sp_pred), ]
tau_pred <- merge(tau_pred, traits[, c("Code", "Mass", "Pitch")],
                  by = "Code")
tau_pred$log_mean <- tau_pred$mean
tau_pred$mean <- exp(tau_pred$mean) * 100
tau_pred$q5 <- exp(tau_pred$q5) * 100
tau_pred$q95 <- exp(tau_pred$q95) * 100
print(tau_pred[, c("Code", "mean", "q5", "q95", "N", "Mass", "Pitch")])
```

Mean EDR for sparrows (except HASP): `r mean(fam_df_sparrow$Tau)`, 
Song Pitch: `r mean(fam_df_sparrow$Pitch)`, 
Mass: `r mean(fam_df_sparrow$Mass)`, 
n = `r nrow(fam_df_sparrow)`

Mean EDR for thrashers (except LCTH): `r mean(fam_df_thrasher$Tau)`, n = `r nrow(fam_df_thrasher)`

Mean EDR for thrushes (except BITH): `r mean(fam_df_thrush$Tau)`, n = `r nrow(fam_df_thrush)`

Mean EDR for warblers (except KIWA): `r mean(fam_df_warbler$Tau)`, n = `r nrow(fam_df_warbler)`